On the Dynamics of Anthropomorphism in Robotics

### 2013-08-15 dynamics-anthropomorphism.tex
Merge of changes operated in HRI version

\abstract{While anthropomorphism in robotics is a commonly discussed trait of
HRI, it paradoxically lacks formal grounds. This article
attempts first at providing a comprehensive synthesis of the social
phenomenon of \textit{perceiving human-like characteristics in non-human
agents and attributing those characteristics to robots}. We draw on
social sciences and psychology, as well as on a critical survey of
existing literature in the HRI community, to ground the concept of
\textit{anthropomorphism}. We then suggest to go beyond the traditional
perception of anthropomorphism as a static feature of a system: we propose to
understand anthropomorphism as a dynamic, non-monotonic and
context-dependent process, which evolves over time and accounts for
special events such as the so-called \textit{novelty effect}. To this end,
we introduce a model of anthropomorphism that analyzes the phenomenon
along three interaction phases. These findings are supported by a long-term
study conducted in a real human environment.}

@@

So far, the HRI community has not much investigated how anthropomorphism in
human-robot interactions evolves over time (during the process of
\emph{adopting} a robot, for instance): we propose to go beyond the traditional
perception of anthropomorphism in robotics as a static feature that once
observed during a short-term interaction reflects a sustaining social effect.
Based on an extensive literature review previously published~\cite{fink_anthropomorphism_2012} and \hl{our own work} \fxfatal{Summarize
these findings in a specific section!}, we believe that anthropomorphic effects in HRI are not always the same but evolve over time, along with growing interaction
and experience with the robot, similar to how relationships among people evolve over time \hl{(cite Hinde, 1988)}. In fact, Kanda \textit{et al.} \cite{kanda_interactive_2004} also hypothesized that people's attitude toward technological artifacts and their relationship with them would evolve over time, and they were among the first who described \textit{novelty effects} during a long-term study with an interactive robot in a school.

The model of anthropomorphism we propose (Figure~\ref{fig:dynamics}) represents
how the level of anthropomorphic effects (\ie observable manifestations of
anthropomorphism, as defined in section~\ref{sec:intro}) evolves over a
long-term human-robot interaction. By long-term interaction, we mean direct
(non-mediated), repeated interaction with the same robot, over at least several
days. The nature of the interaction (goal-directed, entertainment, etc.) may
however vary.

@@

\subsection{Initial Capital of Anthropomorphism}
\label{sec:ica}

A robot has an initial potential of being anthropomorphized. It has been shown that some \textit{people} anthropomorphism more than others, some \textit{situations} induce anthropomorphism more than others, that \textit{children} tend to anthropomorphize more than adults, and some \textit{cultures} are notorious for their anthropomorphic religions and worldviews \cite{epley_when_2008}. Our model of anthropomorphism takes these determinants into account and initializes the level of anthropomorphic interactions between a human and a robot to a value that we call \textit{initial capital of anthropomorphism} (ICA on Figure~\ref{fig:dynamics}). The ICA describes the first (real or imagined) contact to a robot. Time point $t_{0}$ can be temporally before the first real physical interaction with the robot, for instance, when a person learns about the existence of vacuum-cleaning robots and imagines using one (\eg when one watches an advertisement of the robot, or observes one at a trade show). In this stage of "pre-interaction", people form initial expectations toward the robot and imagine how they will use / interact with it. Time point $t_{0}$ can be understood as duration of interaction = 0 and user experience = 0. We build the ICA on three main factors that \textit{a priori} determine the potential that a robot will be anthropomorphized:

@@

    \item \emph{Human-centered factor}: The \textbf{personality} and individual traits of the human user: Psychological characteristics / determinants that influence a person's tendency to anthropomorphize artifacts~\cite{epley_seeing_2007}. Other individual traits and demographic aspects are comprised (\textit{e.g.} age, gender, cultural background, professional background).

@@

    By \textbf{purpose} we suggest that the real or imagined context in which a robot is used and the interaction that it brings along, impacts how far the robot will be attributed human-like characteristics. We	draw on findings such as presented in Joosse \textit{et al.}\fxfatal{reference?}. The authors showed that when the same robot (NAO) is used in a different task context (cleaning task \textit{vs.} tour guide), this influences the perception of the "personality" of the robot. In general, we think that a robot which is imagined to be used in a social, entertaining or playful context leads to a higher ICA than a robot which is used for a serious task (security, rescue, etc.). This idea receives support from \hl{XX (Kiesler / Goetz?)} work that revealed that people prefer a serious robot for serious tasks and a less serious robot for more playful tasks \hl{REF}. Also, we suggest that the environmental context in which people experience and interact with the robot impacts the ICA. When several people who might be friends interact simultaneously with the robot this might lead to an increased ICA \hl{study: people are more emotional when watching TV in company than when watching TV on their own}.

\subsection{Anthropomorphism during Initialization Phase}
\label{sec:initialization}

@@

\subsection{Anthropomorphism during Familiarization Phase}
\label{sec:familiarization}

@@

Our model represents \emph{disruptive behaviors} as local increases of
anthropomorphic effects in the familiarization phase (and at a lesser extend,
in the stabilization phase) (Figure~\ref{fig:dynamics}): because such behaviors
are unexpected, a human observer interprets them as the result of a richer
deliberative process.

@@

\subsection{Anthropomorphism during Stabilization Phase}
\label{sec:stabilization}

If, after the familiarization phase, no anthropomorphic effects are observed
anymore, the \emph{Stabilized Level of Anthropomorphism} is zero. This can either be interpreted as the user
considering the robot as merely a tool, or as a consequence of no interactions
taking place anymore with the robot (robot not used)\fxwarning{reference to the
roomba experiment}.

@@

Note that the \emph{Initial Capital of Anthropomorphism} and the
\emph{Stabilized Level of Anthropomorphism} are generally not correlated: one
individual may have high expectations (high ICA), and get disappointed by the
actual abilities of the robots, down to considering it as a simple machine/tool
(low SLA), while another one with the same high ICA may, for instance, create
lasting affective bonds with the robot, and keep anthropomorphizing it (high
SLA).

@@

\subsection{Anthropomorphism and Cognitive Models}
\label{sec:cognitivemodel}

@@

\end{tikzpicture}
%}
\caption{The three cognitive phases of anthropomorphism: Phase I is the instinctive,
sub-cognitive identification of living peers. {\it Empathy} [TO DISCUSS] is characteristic
of this stage. After longer observation or short, uncontextualized interaction
(typically, a lab environment), the user enters Phase II: the user projects a
mental model he/she is already familiar with onto the robot. This leads to
expectation regarding the cognitive abilities of the robot. After longer {\it
contextualized} interaction (typically, at home), the user enters Phase III of
anthropomorphism: the user recomposes an accurate mental model of the robot,
based on experience. This leads to adapted interaction modalities, that may
still be anthropomorphic, or not.}
\label{fig:cognitivemodel}
\end{figure}

Figure~\ref{fig:cognitivemodel} offers a cognitive perspective on the dynamics
of anthropomorphism presented in the previous section.

The first cognitive phase we identify that contribute to anthropomorphism is
actually \emph{pre-cognitive}. Rosenthal-von der Pütten \textit{et al.} \cite{rosenthal-vonderputten_experimental_2013} investigated the neural correlates of emotional
reactions of human towards robot or other human that showed comparable
responses. \fxfatal{Complete that. Reference to the pleo paper}

After a longer observation period (typically including complete action
sequences of the robot) or short interaction (touching, short talk like
greetings), we suggest the human enters the cognitive \emph{phase II}: in this
phase, the human starts to build a behavioral and cognitive model of the robot
that would support both the observed and imagined capabilities of the robot.
The \emph{familiarity thesis}\fxfatal{citation!} would support the idea that
the human first projects onto the robot mental models of similar agents he/she
is already familiar with. This may range from animals to human adults, through
pets and children. We hypothesize that the nature of the projected mental
model, as well as how deep the human engages in this projection, might be
driven by the same parameters as we presented for the \emph{initial capital of
anthropomorphism} (section~\ref{sec:ica}).

@@

@@

This is explained by the intricacy and interdependency of those behaviors.
Their correct identification and interpretation rely on psychosomatic cues that
the authors did not investigate.

A formal methodology for the assessment of anthropomorphic effects in HRI, as
well as a typology of these effects is still to be devised.

@@

\section{Conclusion}
\label{sec:9}

	Anthropomorphism is a social phenomenon, which seems to be natural on one side and very complex on the other side. It is a mechanism within oneself that makes a human observer think and treat a non-human agent as if it would have some human (social) characteristics, and ascribe to it intentions, emotions, or thoughts, for instance. In accordance with other researchers, we also think that the term is sometimes misused or misunderstood \cite{duffy_anthropomorphism_2002}. Concerning robotics, we do not think that we can speak of anthropomorphism when a person simply refers to physical parts of a robot using terms that describe a human body, such as arm, head, fingers, etc.. We believe that this is not anthropomorphism in its true sense because the person might simply take advantage of the physically resembling shape instead of truly ascribing human (social) characteristics to the robot.

\section*{Acknowledgements}
AJung Moon, Aude Billard.\\

>>>>

\abstract{While anthropomorphism in robotics is a commonly discussed trait of
HRI, it paradoxically lacks formal grounds. This article
attempts first at providing a comprehensive synthesis of the social
phenomenon of \textit{perceiving human-like characteristics in non-human
agents and attributing those characteristics to robots}. We draw on
social sciences and psychology, as well as on a critical survey of
existing literature in the HRI community, to ground the concept of
\textit{anthropomorphism}. Supported by extensive literature review
and long-term field studies, we suggest to go beyond the traditional
perception of anthropomorphism as a static feature of a system: we propose to
understand anthropomorphism as a dynamic, non-monotonic and context-dependent
process, which evolves over time and accounts for special events such as the
so-called \textit{novelty effect}. To this end, we introduce a model of
anthropomorphism that analyzes the phenomenon along three interaction phases,
and we discuss cognitive interpretations that follow.

@@

So far, the HRI community has not much investigated how anthropomorphism in
human-robot interactions evolves over time (during the process of
\emph{adopting} a robot, for instance): we propose to go beyond the traditional
perception of anthropomorphism in robotics as a static feature that once
observed during a short-term interaction reflects a sustaining social effect.
Based on an extensive literature review previously
published~\cite{fink_anthropomorphism_2012} and field
experiements~\fixme{cites!}, we believe that anthropomorphic effects in HRI are
not always the same but evolve over time, along with growing interaction and
experience with the robot, similar to how relationships among people evolve
over time \fixme{(cite Hinde, 1988)}. In fact, Kanda \textit{et al.}
\cite{kanda_interactive_2004} also hypothesized that people's attitude toward
technological artifacts and their relationship with them would evolve over
time, and they were among the first who described \textit{novelty effects}
during a long-term study with an interactive robot in a school.

The model of anthropomorphism we propose (Figure~\ref{fig:dynamics}) represents
how the level of anthropomorphic effects (\ie observable manifestations of
anthropomorphism, as defined in section~\ref{sec:intro}) evolves over a
long-term human-robot interaction. By long-term interaction, we mean direct
(non-mediated), repeated interaction with the same robot, over a long period of
time. The nature of the interaction (goal-directed, entertainment, etc.) may
however vary.

@@

\subsection{Initialization Phase}
\label{sec:initialization}

\paragraph{Initial Capital of Anthropomorphism}
\label{sec:ica}

A robot has an initial potential of being anthropomorphized. It has been shown
that some \textit{people} anthropomorphism more than others, some
\textit{situations} induce anthropomorphism more than others, that
\textit{children} tend to anthropomorphize more than adults, and some
\textit{cultures} are notorious for their anthropomorphic religions and
worldviews \cite{epley_when_2008}. Our model of anthropomorphism takes these
determinants into account and initializes the level of anthropomorphic
interactions between a human and a robot to a value that we call
\textit{initial capital of anthropomorphism} (ICA on
Figure~\ref{fig:dynamics}). The ICA describes the first (real or imagined)
contact to a robot. In this stage of "pre-interaction", people form initial
expectations toward the robot and imagine how they will use / interact with it.

We build the ICA on three main factors that \textit{a priori} determine the
potential that a robot will be anthropomorphized:

@@

    \item \emph{Human-centered factor}: The \textbf{personality} and individual
        traits of the human user: Psychological characteristics / determinants
        that influence a person's tendency to anthropomorphize
        artifacts~\cite{epley_seeing_2007}. Other individual traits and
        demographic aspects are comprised (\eg age, gender, cultural
        background, professional background).

@@

By \textbf{purpose} we suggest that the real or imagined context in which a
robot is used and the interaction that it brings along, impacts how far the
robot will be attributed human-like characteristics. We draw on findings such
as presented in Joosse \textit{et al.}~\cite{joosse_what_2013}. The authors showed
that when the same robot (NAO) is used in a different task context (cleaning
task \textit{vs.} tour guide), this influences the perception of the
"personality" of the robot. In general, we think that a robot which is imagined
to be used in a social, entertaining or playful context leads to a higher ICA
than a robot which is used for a routine or focused task (security, rescue, etc.). This
idea receives support from \fixme{XX (Kiesler / Goetz?)} work that revealed
that people prefer a serious robot for serious tasks and a less serious robot
for more playful tasks \fixme{REF}. Also, we suggest that the environmental
context in which people experience and interact with the robot impacts the ICA.
When several people who might be friends interact simultaneously with the robot
this might lead to an increased ICA \fixme{study: people are more emotional
when watching TV in company than when watching TV on their own}.

@@

TDB

\subsection{Familiarization Phase}
\label{sec:familiarization}

@@

Our model represents \emph{disruptive behaviors} as local increases of
anthropomorphic effects in the familiarization phase (and at a lesser extend,
in the stabilization phase) (Figure~\ref{fig:dynamics}): because such behaviors
are unexpected, a human observer may interpret them as the result of a richer
deliberative process, which in turn leads to the supposition of complex
cognitive skills

@@

\subsection{Stabilization Phase}
\label{sec:stabilization}

The last phase is the \emph{stabilization} phase. The level of anthropomorphic
effects tends to stabilize over a longer time, to reach a \emph{stabilized
level of anthropomorphism} (SLA).

If, after the familiarization phase, no anthropomorphic effects are observed
anymore, the \emph{Stabilized Level of Anthropomorphism} is zero. This can
be interpreted as the user interacting with the robot in a routine way, without
projecting anymore human-like traits on the robot.

@@

Note that the \emph{Initial Capital of Anthropomorphism} and the
\emph{Stabilized Level of Anthropomorphism} are generally not correlated: one
individual may have high potential of anthropomorphizing (high ICA) at first
sight of a good-looking humanoid robot, and get disappointed by the actual
abilities of the robots, down to routine, non-anthropomorphic, interactions
(low SLA), while another one with the same high ICA may, for instance, create
lasting affective bonds with the same robot, and keep anthropomorphizing it
(higher SLA).

@@

\section{Cognitive Interpretations}
\label{sec:cognitivemodel}

\paragraph{Explanations for anthropomorphism}

Anthropomorphism represents just one of many examples of induction whereby
``people reason about an unknown stimulus based on a better-known representation
of a related stimulus"~\cite{epley_when_2008}, in this case reasoning about a
non-human agent based on representation of the self or other humans.

According to Lee \textit{et al.} \cite{lee_human_2005},
there are two main perspectives in explaining people's tendency to
anthropomorphize. The first one explains anthropomorphism from the design of
the artifact (anthropomorphic form in the design). It assumes that humans
directly respond to life-like or social cues that an object or system emits,
without thoughtful mental processing, by simply applying stereotypes and
heuristics to it. In fact, from early childhood on, humans are inherently
well-trained to perceive life \cite{epley_seeing_2007}. Schmitz
\cite{schmitz_concepts_2011} describes that within the visual scope of design,
the outer appearance can have an important impact on the overall perception of
an object. The basic assumption here is that if an artifact appears much like a
human, it is likely to be treated similar to a human. If this explanation of
anthropomorphism is correct, people may respond automatically to social cues
emitted by a robot, and apply human-human social schemas and norms to these
interactions.

The second perspective applies a human-centered, cognitive viewpoint where
anthropomorphism is described through people's specific mental model they
construct about how an artifact works the way it does. According to v.
Foerster \fixme{ref} \textit{``we anthropomorphize because it allows us to
explain things we do not understand in terms that we do understand, and what we
understand best is ourselves as human beings"}~\cite{hegel_understanding_2008}.
This is consistent with the \emph{familiarity
thesis}~\cite{hegel_understanding_2008} which claims we understand the world
based upon a mental model of the world that we are most familiar with.
Consequently, people tend to thoughtfully develop a mental model of agents in
their environment and make inferences about it based on what is familiar to
them. This point of view implicitly builds on a person's ability to attribute
mental states to oneself and others (\ie the availability of a \emph{theory of
mind}~\cite{Premack1978a}). A theory of  mind for other agents enables us to
attribute intentionality to those agents
\cite{leslie_pretense_1987,admoni_multi-category_2012}. Previous research
examined the validity of the mental model concept with various kinds of robots
\cite{schmitz_concepts_2011,kiesler_mental_2002}. Findings suggest that people
tend to hold richer mental models about anthropomorphic robots in contrast to
mechanic ones \cite{kiesler_mental_2002}.

@@

\end{tikzpicture}
%}
\caption{The three cognitive phases of anthropomorphism: Phase I is the instinctive,
sub-cognitive identification of living peers. {\it Empathy} is characteristic
of this stage. After longer observation or short, uncontextualized interaction
(typically, a lab environment), the user enters Phase II: the user projects a
mental model he/she is already familiar with onto the robot. After longer {\it
contextualized} interaction (typically, at home), the user enters Phase III of
anthropomorphism: the user recomposes an accurate mental model of the robot,
based on experience. This leads to adapted interaction modalities, that may
still be anthropomorphic, or not.}
\label{fig:cognitivemodel}
\end{figure}

\paragraph{Building cognitive models over time}

Building on these previous researches, Figure~\ref{fig:cognitivemodel} offers a
cognitive perspective on the dynamics of anthropomorphism presented in the
previous section.

The first cognitive phase we identify that contribute to anthropomorphism is
actually \emph{pre-cognitive}. This is supported by studies like Rosenthal-von
der Pütten \textit{et al.}~\cite{rosenthal-vonderputten_experimental_2013} that
investigated the neural correlates of emotional reactions of human towards
robot.

After a longer observation period (typically including complete action
sequences of the robot) or short interaction (touching, short talk like
greetings), we suggest the human enters the cognitive \emph{phase II}: in this
phase, the human starts to build a behavioral and cognitive model of the robot
that would support both the observed and imagined capabilities of the robot.
The \emph{familiarity thesis} would support the idea that
the human first projects onto the robot mental models of similar agents he/she
is already familiar with (ranging from animals to human adults, through
pets and children). We hypothesize that the nature of the projected mental
model, as well as how deep the human engages in this projection, might be
driven by the same parameters as we presented for the \emph{initial capital of
anthropomorphism} (section~\ref{sec:ica}).

@@

\subsection{Impact of the behavioral complexity of robots}

TBD. Most experiments are carried with robots exhibiting relatively simple
behaviors. In our model, the maximum level of anthropomorphism is reached with
the novelty effect, and then decrease. Can we go \emph{over} the novelty effect
if the robot exhibit more complex behaviors?

\subsection{Segmentation of interaction phases}

TBD

The intricacy and interdependency of those behaviors make them difficult to
quantify.  Their systematic identification and interpretation rely on
psychosomatic cues that the authors did not investigate.

A formal methodology for the assessment of anthropomorphic effects in HRI, as
well as a typology of these effects is still to be formally devised.

@@

@@

\section{Conclusion}
\label{sec:conclusion}

Anthropomorphism is a social phenomenon which seems to be natural on one side
and very complex on the other side. It is a mechanism within oneself that makes
a human observer think and treat a non-human agent as if it would have some
human (social) characteristics, and ascribe to it intentions, emotions, or
thoughts, for instance. Understanding the reasons and modalities of
anthropomorphism is thus of prime importance to build a successful and lasting
human-robot interaction.

Anthropomorphism is traditionally empirically understood as the combination of
the anthropomorphic design of a robot and the psychological determinants of the
user. We have found out that the duration and context of the interaction is a
third factor that plays a key role. In this article, we propose a new formal
model of anthropomorphism that accounts for these three factors and also
explicits the dynamics of anthromorphism. We introduce the concepts of
\emph{initial} and \emph{stabilized levels of anthropomorphism} as compound
factors to characterize the profile of a given anthropomorphic interaction.

The article also discusses the cognitive correlates of anthropomorphism. We
proposes to identify three cognitive phases corresponding to successive
refinements of the mental models of the robot that the user builds during the
interaction. We show how these phases relate to observable anthorpomorphic
behaviours, and how they evolve over time.

While not definitive, we hope that this contribution consolidates the
scientific grounds of anthropomorphism, and provides support for further
research on long-term acceptance of robots in human environments.

\section*{Acknowledgements}

We would like to acknowledge here the fruitful discussions with A Jung Moon and
Aude Billard.\\
